---
title: "Section 4: Intervals, PPCS, and Rejection Sampling"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```


## Hot Hand
The "hot hand" is the purported phenomenon that a person who experiences a successful outcome has a greater chance of success in further attempts. The concept is originates from basketball whereas a shooter is allegedly more likely to score if their previous attempts were successful. While previous success at a task can indeed change the psychological attitude and subsequent success rate of a player, researchers for many years did not find evidence for a "hot hand" in practice, dismissing it as fallacious. However, later research questioned whether the belief is indeed a fallacy. In this lab, we will boil down this qusetion as testing whether the results of shots are independent.

Let "1” denotes a valid shot and "0" denotes a invalid. Suppose we observe the following results of a player:
```{r}
# observations #
set.seed(123)
y <- c(rep(1, 18), rep(0, 3), rep(1, 6), rep(0, 2),
       rbinom(67, 1, prob = 0.25), rep(1, 4))
```

Suppose $Y_i \sim Bernoulli(\theta)$ and $\theta \sim Beta(3, 7)$

Find the posterior:
```{r}
# prior #
a <- 3
b <- 7
#posterior #
a_post <- a + sum(y)
b_post <- b + (length(y) - sum(y))
a_post; b_post
```

Let the test stat. be the maximum number of the same consecutive results. (Other good test stat. might be the fraction of times you see (1, 1) vs (1, 0).)
```{r, warning = F, message = F}
# observed test stat. #
test_stat_obs <- max(rle(y)$lengths)

# test stat. based on simulation #
nsim <- 1000
test_stat_rep <- rep(NA, nsim)
for (i in 1:1000) {
  p_post <- rbeta(1, a_post, b_post)
  y_rep <- rbinom(100, size = 1, prob = p_post)
  test_stat <- max(rle(y_rep)$lengths)
  test_stat_rep[i] <- test_stat
}

ggplot(tibble(test_stat_rep), aes(test_stat_rep)) + 
  geom_histogram() + xlab("Max Num.") + 
  geom_vline(xintercept = test_stat_obs, colour = "red")
```

In this example, the data doesn't look independent. The "hot-hand" phenomenon might exist. 













In this section we'll review rejection sampling.

## Rejection Sampling

Rejection sampling is easiest to think of graphically. We first plot the probability density function or probability mass function of the distribution that we want to sample from. If we can uniformly sample from a region that bounds this pdf/pmf, then we can do rejection sampling by:

#. Rejection Sampling for the Semi-Circle Distribution.  Let 
    $$p(x|R)=\frac{2}{\pi R^2}\sqrt{R^2-x^2}I\{x \in [-R, R]\}$$
 
    <!-- solution end -->

    
    #. Identify a proposal distribution, $g(x \mid R)$ for rejection sampling.  Show that there exists some $M$ such that $Mg(x \mid R) \geq p(x|R)$ $\forall x \in [-R, R]$.  What is M? 
    
     <!-- solution begin -->
        
    proposal distribution: $$g(x|R) = \frac{1}{2R}\mathbf{1}\{x \in [-R, R]\}$$
    \begin{align*}
    \frac{p(x|R)}{g(x|R)} &= \frac{4}{\pi R}\sqrt{R^2 - x^2} \\
    & \leq \frac{4}{\pi R} R\\
    &=\frac{4}{\pi} 
    \end{align*}
    Therefore, M can choose $\frac{4}{\pi}.$
    
    <!-- solution end -->
    
    #.   Plot the semi-circle distribution for $R=1$ and your proposal distribution on the same plot using the `curve` function.  Note that because the density must be normalized to integrate to 1, the semi-circle distribution is really a "semi-ellipse". 
    <!-- solution begin -->
```{r}

ratio_densities <- function(x) {
  (2/pi*sqrt(1-x^2)) / (2*dbeta((x + 1)/2, 1.2, 1.2))
}

M <- optimize(ratio_densities, lower=-1, upper=1, maximum=TRUE)$objective

curve(2/pi*sqrt(1-x^2), from=-1, 1, ylim=c(0, 1))
curve(M*2*dbeta((x+1)/2, 1.2, 1.2), add=TRUE, col="blue")
abline(h=2/pi, col="red")
```
    
    
    
    #.  Use rejection sampling to sample from the semi-circle distribution with $R=1$.  What is the acceptance rate of your  sampler?

  <!-- solution begin -->

    
```{r}
N = 10000
R = 1
M = 4/pi
x <- runif(N, -R, R)
y <- runif(N, 0, 1)
p <- function(x, R) {
  return(2/pi/R^2 * sqrt(R^2 - x^2))
}
g <- function(x, R) {
  return(1/2/R)
}
decision <- y <= sapply(x, p, R = 1)/M /sapply(x, g, R = 1)

x = x[decision]
sum(decision)/ N
```

  <!-- solution end -->
    #. For $R = 0.1, 0.2, ..., 3.0$, use rejection sampling to sample from $p(x|R)$.  Plot the sample variance for each value of $R$, and show graphically that the variance for the semi-circle distribution is $\frac{R^2}{4}$.
  <!-- solution begin -->
```{r}
R = seq(0.1, 3.0, by = 0.1)
M = 4/pi
N = 10000
v <- rep(0, length(R))
for (i in 1 : length(R)) {
  r = R[i]
  x <- runif(N, -r, r)
  y <- runif(N, 0, 1)
  decision <- y <= sapply(x, p, R = r)/M /sapply(x, g, R = r)
  x = x[decision]
  v[i] = var(x)
}
v <- v %>% as.data.frame() %>% cbind(R)
colnames(v) <- c("var", "R")
ggplot(v, aes(x = R, y = var)) + geom_line()

```
  It can be checked that the line resembles $var = R^2/4.$

  <!-- solution end -->

```{r}
N = 1000

alpha = 2.0
beta = 3.0
mode = (alpha - 1) / (alpha + beta - 2)
maxB = dbeta(mode, alpha, beta)

list(x = runif(N),
     y = runif(N) * maxB) %>%
  as.tibble %>%
  mutate(use = y < dbeta(x, alpha, beta)) %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = use)) +
  stat_function(fun = function(x) dbeta(x, alpha, beta)) +
  geom_rect(aes(xmin = 0.0, xmax = 1.0, ymin = 0.0, ymax = maxB), alpha = 0.0, color = "red") +
  ggtitle("Proposal distribution outlined in red, target in black")
```

The points highlighted in teal above have the desired distribution. We throw away the reddish samples.